---
title: "BusDelay_Logistic_Regression"
author: "tonghua.lin@rutgers.edu"
date: "4/19/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## R Markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In last assignment, I found that amount the different month, May seemed important for delays. Thus, for logistic regression, I want to check whether we can divide May from other months by data.
```{r}
# Read the data to DataSet
library(readr)
library(ggplot2)
library(cowplot)
library(lattice)
library(caret)
library(e1071)
library(pROC)
DataSet <- read_csv("./BUS_OTP_DATA.csv")
print(DataSet)

```

Assume the total lates will change stably by years, here we have years, OTP and total trips as numerical variables. For month, we need to set dummy variables with n-1 rule. Assume January is the base.
```{r}
# Created dummies for month.
DataSet$OTP_MONTH <- ifelse(test=DataSet$OTP_MONTH == "May", yes="May", no="OtherMonth") 
DataSet$OTP_MONTH <- as.factor(DataSet$OTP_MONTH)

str(DataSet)
```
Let us check the relationship between TOTAL_LATES and OTP_MONTH first.
```{r}
# perform the logistic regression 
logistic_simple <- glm(OTP_MONTH ~ TOTAL_LATES, data=DataSet, family="binomial")
summary(logistic_simple)
```
The results seem not very well. Let check other variables.
```{r}
logistic <- glm(OTP_MONTH ~ ., data=DataSet, family="binomial")
summary(logistic)
```
The results show that with other variables, TOTAL_TRIPS and TOTAL_LATES has weakly significant influence. Namely if we know the TOTAL_TRIPS and TOTAL_LATES of a month, we may know whether it's May. Following is some visualization.
```{r}
predicted <- data.frame(probabilityMay=logistic$fitted.values,OTP_MONTH=DataSet$OTP_MONTH)
predicted <- predicted[order(predicted$probabilityMay, decreasing=FALSE),]
predicted$rank <- 1:nrow(predicted)
ggplot(data=predicted, aes(x=rank, y=probabilityMay)) +
geom_point(aes(color=OTP_MONTH), alpha=1, shape=4, stroke=2) +
xlab("Index") +
ylab("Predicted probability of May")
```
