---
title: "Social_Multiple_Regression"
author: "tonghua.lin@rutgers.edu"
date: "4/14/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## R Markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

For this data set, I want to set Weekfeeling as the output, and see which variables will significantly influence the output. The independent varibles include different apps, weekly results etc.
```{r}
library(readr)
DataSet <- read_csv("Social Media.csv")
print(DataSet)
DataSet$'Application type(Social media, OTT, Learning)'<-NULL
#drop the string column and all APP Yes/No column since time=0 means No
DataSet$Linkedin <- NULL
DataSet$Whatsapp <- NULL
DataSet$ApplicationType <- NULL
DataSet$Instagram <- NULL
DataSet$Snapchat <- NULL
DataSet$Twitter <- NULL
DataSet$Youtube <- NULL
DataSet$OTT <- NULL
DataSet$Reddit <-NULL

#change the unit of hms variables to minutes
hms_columns <- sapply(DataSet, inherits, "hms")
DataSet[hms_columns] <- lapply(DataSet[hms_columns], function(x) as.numeric(x) / 60)
print(DataSet)
```
With the processed data, we can start a multiple regression.
```{r}
# Performing multiple regression
fit <- lm(WeekFelt~InsTime+LinkedinTime+SnapTime+TwitterTime+WhatsappTime+YoutubeTime+OTTTime+RedditTime+JobInterview+Network+LearningItem+TiredWakeUP+TroubleSleep, data=DataSet)
#show the results
summary(fit)
coefficients(fit)
```
From the result, we can see that there are many insignificant variables. Here, we can choose significant variables to rerun the regression model to get better results.
```{r}
fit <- lm(WeekFelt~InsTime+LinkedinTime+TwitterTime+WhatsappTime+OTTTime+JobInterview+LearningItem+TiredWakeUP, data=DataSet)
#show the results
summary(fit)
coefficients(fit)
```
With the result, we can see the R-squared value of the model is 0.70, which is a acceptable result. Also, most of remaining variables are significant. Now, we can run a residual analysis.
```{r}
#start a residual analysis
confint(fit,level=0.95)
fitted(fit)
residuals(fit)
#Anova Table
anova(fit)
vcov(fit)
cov2cor(vcov(fit))
temp <- influence.measures(fit)
temp
plot(fit)
```
From the results, we can see the residual distribution is reasonable. Again, this proves the validation of the model. We can use the model to predict an output if we have data.
```{r}
predict.lm(fit, data.frame(InsTime=300,LinkedinTime=150,TwitterTime=200,WhatsappTime=300,OTTTime=60,JobInterview=1,LearningItem=3,TiredWakeUP=0))
```

