---
title: "Social_Logistic_Regression"
author: "tonghua.lin@rutgers.edu"
date: "4/14/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## R Markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

For this data set, I want to check the relationship between trouble sleep with other variables. We need to clean the data first.

```{r}
library(readr)
DataSet <- read_csv("Social Media.csv")
DataSet$'Application type(Social media, OTT, Learning)'<-NULL
#drop the string column and all APP Yes/No column since time=0 means No
DataSet$Linkedin <- NULL
DataSet$Whatsapp <- NULL
DataSet$ApplicationType <- NULL
DataSet$Instagram <- NULL
DataSet$Snapchat <- NULL
DataSet$Twitter <- NULL
DataSet$Youtube <- NULL
DataSet$OTT <- NULL
DataSet$Reddit <-NULL
DataSet$ID <- NULL
DataSet$Mood <- NULL
DataSet$WeekFelt <- NULL

#change the unit of hms variables to minutes
hms_columns <- sapply(DataSet, inherits, "hms")
DataSet[hms_columns] <- lapply(DataSet[hms_columns], function(x) as.numeric(x) / 60)
print(DataSet)
```
Change JobInterview, Network, Learning Item, Mood, TiredWakeUp and WeekFelt to factors.
```{r}


DataSet$TroubleSleep <- as.factor(DataSet$TroubleSleep)
str(DataSet)
```
TirdeWakeUp may be important for troubleSleep intuitively. We can check with a simple regression.
```{r}
logistic_simple <- glm(TroubleSleep ~ TiredWakeUP, data=DataSet, family="binomial")
summary(logistic_simple)
```
Not very significant. Let's check other variables.
```{r}
logistic <- glm(TroubleSleep ~ ., data=DataSet, family="binomial")
summary(logistic)
```
The results show that DataSet is too small to get some logistic regression model.

