---
title: "BusDelay_Multiple_Regression"
author: "tonghua.lin@rutgers.edu"
date: "4/14/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## R Markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

For this data set, I want to set total lates as th output, and see which variables will significantly influence the output. The independent varibles are years, month, OTP and total trips.
```{r}
# Read the data to DataSet
library(readr)
DataSet <- read_csv("./BUS_OTP_DATA.csv")
print(DataSet)

```

Assume the total lates will change stably by years, here we have years, OTP and total trips as numerical variables. For month, we need to set dummy variables with n-1 rule. Assume January is the base.
```{r}
# Created dummies for month.
DataSet$OTP_MONTH <- factor(DataSet$OTP_MONTH, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
dummies <- model.matrix(~ OTP_MONTH, data = DataSet)
dummies <- dummies[, -1]
colnames(dummies) <- month.abb[-1]
DataSetDum <- cbind(DataSet, dummies)
DataSetDum$OTP_MONTH<-NULL
print(DataSetDum)

```
With the processed data, we can start a multiple regression.
```{r}
# Performing multiple regression
fit <- lm(TOTAL_LATES~OTP_YEAR+OTP+TOTAL_TRIPS+Feb+Mar+Apr+May+Jun+Jul+Aug+Sep+Oct+Nov+Dec, data=DataSetDum)
#show the results
summary(fit)
coefficients(fit)
```
From the result, we can see that there are many insignificant variables. Here, we can choose significant variables to rerun the regression model to get better results.
```{r}
fit <- lm(TOTAL_LATES~OTP+TOTAL_TRIPS+May, data=DataSetDum)
#show the results
summary(fit)
coefficients(fit)
```
With the result, we can see the R-squared value of the model is 0.98, which is a highly acceptable result. Also, remaining variables are all significant. Now, we can run a residual analysis.
```{r}
#start a residual analysis
library(ggplot2)
library(GGally)
ggpairs(data=DataSetDum, title="Cars Data")
confint(fit,level=0.95)
fitted(fit)
residuals(fit)
#Anova Table
anova(fit)
vcov(fit)
cov2cor(vcov(fit))
temp <- influence.measures(fit)
temp
plot(fit)
```
From the results, we can see the residual distribution is quite reasonable. Again, this proves the validation of the model. We can use the model to predict an output if we have data.
```{r}
predict.lm(fit, data.frame(OTP=95 ,TOTAL_TRIPS=32000,May=0))
```
Also, we can divide the data to two part and use the prediction result to check the accuracy of the model.
```{r}
# Randomize rows
set.seed(123)  # Setting a seed to make the randomization reproducible
randomized <- DataSetDum[sample(nrow(DataSetDum)), ][,c("OTP","TOTAL_TRIPS","TOTAL_LATES","May")]
# Reset row names to avoid confusion
row.names(randomized) <- NULL
print(randomized)
data_train=randomized[1:150,]
data_test=randomized[151:180,]
fit2 <- lm(TOTAL_LATES~OTP+TOTAL_TRIPS+May, data=data_train)
summary(fit2)
predict_lates <- predict.lm(fit2,data_test)
difference = predict_lates - data_test$TOTAL_LATES
accuracy <- 1 - mean(abs(difference))/mean(data_test$TOTAL_LATES)
print(accuracy)
```
We can see the accuracy of the model can be 98%.
